|

### 2020-11-30 manuscript/article.Rmd
commit 1 11/30

The binomial z-test evaluates whether the coefficient of our independent
variables are positive or negative. The null hypothesis here states that the
sign of the coefficient is equally likely to be positive or negative. We start
with $n$, lower house size. Our results indicate that there is no correlation
between the number of legislators in the lower house and public expenditure 
(successes = 10, trials = 21, $p_{success} = 0.476$, 95\% CI = [0.257; 0.702],
$p$-value = 1). Note that the "law of $1/n$" suggests that there is a positive
association between both. The binomial test for log($n$) also shows a
non-statistically significant result (successes = 5, trials = 6, $p_{success} =
0.833$, 95\% CI = [0.359; 0.996], $p$-value = 0.219). In contrast, we find a
positive result for the number of legislators in the upper house ($k$), which is
in line with the mainstream literature (successes = 8, trials = 9, $p_{success}
= 0.889$, 95\% CI = [0.518; 0.997], $p$-value = 0.039).

@@

Figure \ref{fig:plots} shows the forest plots for our restricted sample, the one
which includes only the main 36 coefficients for the 26 selected papers. On the
left side of the plots are the name of the study authors and the paper
publication year. For unpublished studies, we included the first year the paper
was available online. The bars in the middle show the reported effect sizes and
the vertical lines indicate their average, weighted by standard errors. The length
of the lines represent the precision of the estimates. The red line at the
bottom of the figures displays the aggregate coefficients plus their respective
confidence intervals.

The first column of the graph reports the results for $n$, lower chamber size.
In the first model, which correlates $n$ and expenditure per capita, we find a
standardised mean difference (SMD) of $-0.070$ and a standard error of $0.273$
(studies = 12, 95\% CI = [-0.671; 0.531], $p$-value = 0.803, prediction interval
= [-1.55; 1.41]). Therefore, we cannot reject the null hypothesis that the
effect size is zero. The $I^2$ statistic, which quantifies the degree of
heterogeneity among studies, is equal to 94.7\%. $I^2$ values above 75\%
indicate very high study heterogeneity \citep{higgins2019cochrane}.

We see similar results when we compare $n$ with log expenditure per capita. In
this model, the average effect size is -0.07 and the standard deviation is
0.059 (studies = 4, 95\% CI = [-0.256; 0.119], $p$-value = 0.328, prediction
interval = [-0.618; 0.481]; $I^2$ = 92.5\%). The coefficient is 
statistically indistinguishable from zero.

In our restricted sample, larger lower house size also does not increase
government spending as a percentage of GDP (studies = 5, SMD = -0.008, SE =
0.013, 95\% CI = [-0.045; 0.0285], $p$-value = 0.567, prediction interval =
[-0.105; 0.089], $I^2$ = 87.1\%). We find a positive effect in our extended
sample and the coefficient is statistically significant at the 10\% level
(coefficients = 21, SMD = 0.008, SE = 0.004, 95\% CI = [-0.000; 0.016], $p$-value =
0.058, prediction interval = [-0.026; 0.042], $I^2$ = 98.5\%). The 
result supports the "law of 1/$n$", but the evidence is moderate
as the prediction interval include zero.

@@

Next, we present the meta-analyses using log($n$), the logarithm of lower house
size, as the main explanatory variable. We start with the relationship between
log($n$) and the logarithm of expenditure per capita. The result is positive and
statistically significant at 10\%, but the prediction interval encompasses zero
(studies = 3, SMD = 0.184, SE = 0.06, 95\% CI = [-0.0738; 0.4425], $p$-value =
0.0916, prediction interval = [-1.258; 1.627], $I^2$ = 85.9\%). Although the
coefficient has the predicted sign, we should interpret the finding with caution
as it does not replicate in our sample of 126 effect sizes.

Our model that correlates log($n$) with public expenditures as a percentage of
GDP fails to reach conventional levels of statistical significance (studies =
3, SMD = 0.0203, SE = 1.677, 95\% CI = [-7.196; 7.237], $p$-value = 0.991,
prediction interval = [-36.206; 36.246], $I^2$ = 96.1\%). The extended sample
also gives us a null result.

The third set of models uses $k$, upper house size, as the main independent
variable. We find a positive correlation between $k$ and expenditure per capita
and the coefficient is significant at a 10\% level (studies = 6, SMD = 10.613, SE
= 5.148, 95\% CI = [-2.621; 23.848], $p$-value = 0.094, prediction interval =
[-21.130; 42.357], $I^2$ = 79.4\%). But as with the other models, the prediction
interval again includes zero. When we run the same analysis in the extended
sample, we also see a significant coefficient yet a prediction interval that
contains zero (coefficients = 24, SMD = 7.216, SE = 1.342, 95\% CI = [4.440; 9.992],
$p$-value $<$ 0.0001, prediction interval = [-1.222; 15.654], $I^2$ = 77.7\%).

Our last estimation analyses the relationship between $k$ and government
spending as a percentage of GDP. The coefficient is not statistically
significant, indicating a null effect (studies = 3, SMD = -0.003, SE =, 95\% CI
= [-0.079; 0.074], $p$-value = 0.891, prediction interval = [-0.428; 0.423],
$I^2$ = 85.8\%). The result is very similar in the extended sample.

@@

In this article, we assess the empirical validity of the "law of $1/n$". Based
on a sample of 26 recent publications on the topic, our meta-analyses show that
there is little evidence that an increase in the number of legislators has a
significant effect on public expenditures. If such effect exists, it is likely
driven by an increase in $k$, the size of the upper legislature, as suggested
by several studies in the literature \citep[e.g.,][]{baqir2002districting,
bradbury2001legislative, bradbury2003local, chen_malhotra_2007,
gilligan2001fiscal, primo2006stop}. We find no robust evidence for the "reverse
law of $1/n$", which posits that larger legislatures lead to lower government
spending. One possible explanation is that the two logics are balancing each
other, thus leading us to a null impact of legislature size on public
expenditure.

>>>>

```{r, cache=TRUE, echo = FALSE}
aux <- filter(dat, indepvar2 == "N")
binN <- binom.test(table(aux$scoef)[2], sum(table(aux$scoef)), p = 0.5)
aux <- filter(dat, indepvar2=='K')
binK <- binom.test(table(aux$scoef)[2], sum(table(aux$scoef)), p=0.5)
aux <- filter(dat, indepvar2 == "logN")
binlogN <- binom.test(table(aux$scoef)[2], sum(table(aux$scoef)), p = 0.5)
```

The binomial z-test evaluates whether the coefficient of our independent
variables are positive or negative. The null hypothesis here states that the
sign of the coefficient is equally likely to be positive or negative. We start
with $n$, lower house size. Our results indicate that there is no correlation
between the number of legislators in the lower house and public expenditure 
(successes = `r as.numeric(binN$statistic)`, trials = `r as.numeric(binN$parameter)`, $p_{success} = $ `r round(as.numeric(binN$estimate), 3)`, 95\% CI = [`r round(as.numeric(binN$conf.int[1]), 3)`; `r round(as.numeric(binN$conf.int[2]), 3)`],
$p$-value = `r round(as.numeric(binN$p.value), 3)`). Note that the "law of $1/n$" suggests that there is a positive
association between both. The binomial test for log($n$) also shows a
non-statistically significant result (successes = `r as.numeric(binlogN$statistic)`, trials = `r as.numeric(binlogN$parameter)`, $p_{success} = $ `r round(as.numeric(binlogN$estimate), 3)`, 95\% CI = [`r round(as.numeric(binlogN$conf.int[1]), 3)`; `r round(as.numeric(binlogN$conf.int[2]), 3)`],
$p$-value = `r round(as.numeric(binlogN$p.value), 3)`). In contrast, we find a
positive result for the number of legislators in the upper house ($k$), which is
in line with the mainstream literature (successes = `r as.numeric(binK$statistic)`, trials = `r as.numeric(binK$parameter)`, $p_{success} = $ `r round(as.numeric(binK$estimate), 3)`, 95\% CI = [`r round(as.numeric(binK$conf.int[1]), 3)`; `r round(as.numeric(binK$conf.int[2]), 3)`],
$p$-value = `r round(as.numeric(binK$p.value), 3)`).

@@

Figure \ref{fig:plots} shows the forest plots for our restricted sample, the one
which includes only the main `r dim(dat)[1]` coefficients for the `r length(unique(dat$id))` selected papers. On the
left side of the plots are the name of the study authors and the paper
publication year. For unpublished studies, we included the first year the paper
was available online. The bars in the middle show the reported effect sizes and
the vertical lines indicate their average, weighted by standard errors. The length
of the lines represent the precision of the estimates. The red line at the
bottom of the figures displays the aggregate coefficients plus their respective
confidence intervals.

```{r, echo = FALSE, message = FALSE, warning = FALSE, cache=TRUE}
mod <- list()
# Pooling effects analysis -- ExpPC x N
aux <- dat %>%
  filter(indepvar2 == 'N',
         depvar2 == 'ExpPC')

mod[[1]] <- metagen(coef, SE, data=aux,
          studlab=paste(authoryear),
          comb.fixed = FALSE,
          comb.random = TRUE,
          method.tau = "REML",
          hakn = TRUE,
          prediction=TRUE,
          sm="SMD")

# Pooling effects analysis -- ExpPC x K
aux <- dat %>%
  filter(indepvar2 == 'K',
         depvar2 == 'ExpPC')

mod[[2]] <- metagen(coef, SE, data=aux,
          studlab=paste(authoryear),
          comb.fixed = FALSE,
          comb.random = TRUE,
          method.tau = "REML",
          hakn = TRUE,
          prediction=TRUE,
          sm="SMD")
# Pooling effects analysis -- logExpPC x N
aux <- dat %>%
  filter(indepvar2 == 'N',
         depvar2 == 'logExpPC')

mod[[3]] <- metagen(
  coef, SE, data=aux,
  studlab=paste(authoryear),
  comb.fixed = FALSE,
  comb.random = TRUE,
  method.tau = "REML",
  hakn = TRUE,
  prediction = TRUE,
  sm="SMD"
  )

# Pooling effects analysis -- logExpPC x logN
aux <- dat %>%
  filter(indepvar2 == 'logN',
         depvar2 == 'logExpPC')

mod[[4]] <- metagen(coef, SE, data=aux,
          studlab=paste(authoryear),
          comb.fixed = FALSE,
          comb.random = TRUE,
          method.tau = "REML",
          hakn = TRUE,
          prediction=TRUE,
          sm="SMD")

# Pooling effects analysis -- PCTGDP x N
aux <- dat %>%
  filter(indepvar2 == 'N',
         depvar2 == 'PCTGDP')

mod[[5]] <- metagen(coef, SE, data=aux,
          studlab=paste(authoryear),
          comb.fixed = FALSE,
          comb.random = TRUE,
          method.tau = "REML",
          hakn = TRUE,
          prediction=TRUE,
          sm="SMD")

# Pooling effects analysis -- PCTGDP x logN
aux <- dat %>%
  filter(indepvar2 == 'logN',
         depvar2 == 'PCTGDP')

mod[[6]] <- metagen(
  coef, SE, data=aux,
  studlab=paste(authoryear),
  comb.fixed = FALSE,
  comb.random = TRUE,
  method.tau = "REML",
  hakn = TRUE,
  prediction=TRUE,
  sm="SMD"
  )

# Pooling effects analysis -- PCTGDP x K
aux <- dat %>%
  filter(indepvar2 == 'K',
         depvar2 == 'PCTGDP')

mod[[7]] <- metagen(coef, SE, data=aux,
          studlab=paste(authoryear),
          comb.fixed = FALSE,
          comb.random = TRUE,
          method.tau = "REML",
          hakn = TRUE,
          prediction=TRUE,
          sm="SMD")

```

The first column of the graph reports the results for $n$, lower chamber size.
In the first model, which correlates $n$ and expenditure per capita, we find a
standardised mean difference (SMD) of `r round(mod[[1]]$TE.random,3)` and a standard error of `r round(mod[[1]]$seTE.random,3)`
(studies = `r mod[[1]]$k`, 95\% CI = [`r round(mod[[1]]$lower.random,3)`; `r round(mod[[1]]$upper.random,3)`], $p$-value = `r round(mod[[1]]$pval.random,3)` , prediction interval
= [`r round(mod[[1]]$lower.predict,3)`; `r round(mod[[1]]$upper.predict,3)`]). Therefore, we cannot reject the null hypothesis that the
effect size is zero. The $I^2$ statistic, which quantifies the degree of
heterogeneity among studies, is equal to `r round(mod[[1]]$I2*100, 2)`\%. $I^2$ values above 75\%
indicate very high study heterogeneity \citep{higgins2019cochrane}.

We see similar results when we compare $n$ with log expenditure per capita. In
this model, the average effect size is `r round(mod[[3]]$TE.random,3)` and the standard error is
`r round(mod[[3]]$seTE.random,3)` (studies = `r mod[[3]]$k`, 95\% CI = [`r round(mod[[3]]$lower.random,3)`; `r round(mod[[3]]$upper.random,3)`], $p$-value = `r round(mod[[3]]$pval.random,3)`, prediction
interval = [-0.618; 0.481]; $I^2$ = 92.5\%). The coefficient is 
statistically indistinguishable from zero.

In our restricted sample, larger lower house size also does not increase
government spending as a percentage of GDP (studies = `r mod[[5]]$k`, SMD = `r round(mod[[5]]$TE.random,3)`, SE =
`r round(mod[[5]]$seTE.random,3)`, 95\% CI = [`r round(mod[[5]]$lower.random,3)`; `r round(mod[[5]]$upper.random,3)`], $p$-value = `r round(mod[[5]]$pval.random,3)`, prediction interval =
[-0.105; 0.089], $I^2$ = 87.1\%). We find a positive effect in our extended
sample and the coefficient is statistically significant at the 10\% level
(coefficients = 21, SMD = 0.008, SE = 0.004, 95\% CI = [-0.000; 0.016], $p$-value =
0.058, prediction interval = [-0.026; 0.042], $I^2$ = 98.5\%). The 
result supports the "law of 1/$n$", but the evidence is moderate
as the prediction interval include zero.

@@

Next, we present the meta-analyses using log($n$), the logarithm of lower house
size, as the main explanatory variable. We start with the relationship between
log($n$) and the logarithm of expenditure per capita. The result is positive and
statistically significant at 10\%, but the prediction interval encompasses zero
(studies = `r mod[[4]]$k`, SMD = `r round(mod[[4]]$TE.random,3)`, SE = `r round(mod[[4]]$seTE.random,3)`, 95\% CI = [`r round(mod[[4]]$lower.random,3)`; `r round(mod[[4]]$upper.random,3)`], $p$-value =
`r round(mod[[4]]$pval.random,3)`, prediction interval = [-1.258; 1.627], $I^2$ = 85.9\%). Although the
coefficient has the predicted sign, we should interpret the finding with caution
as it does not replicate in our sample of 126 effect sizes.

Our model that correlates log($n$) with public expenditures as a percentage of
GDP fails to reach conventional levels of statistical significance (studies =
`r mod[[6]]$k`, SMD = `r round(mod[[6]]$TE.random,3)`, SE = `r round(mod[[6]]$seTE.random,3)`, 95\% CI = [`r round(mod[[6]]$lower.random,3)`; `r round(mod[[6]]$upper.random,3)`], $p$-value = `r round(mod[[6]]$pval.random,3)`,
prediction interval = [-36.206; 36.246], $I^2$ = 96.1\%). The extended sample
also gives us a null result.

The third set of models uses $k$, upper house size, as the main independent
variable. We find a positive correlation between $k$ and expenditure per capita
and the coefficient is significant at a 10\% level (studies = `r mod[[2]]$k`, SMD = `r round(mod[[2]]$TE.random,3)`, SE
= `r round(mod[[2]]$seTE.random,3)`, 95\% CI = [`r round(mod[[2]]$lower.random,3)`; `r round(mod[[2]]$upper.random,3)`], $p$-value = `r round(mod[[2]]$pval.random,3)`, prediction interval =
[-21.130; 42.357], $I^2$ = 79.4\%). But as with the other models, the prediction
interval again includes zero. When we run the same analysis in the extended
sample, we also see a significant coefficient yet a prediction interval that
contains zero (coefficients = 24, SMD = 7.216, SE = 1.342, 95\% CI = [4.440; 9.992],
$p$-value $<$ 0.0001, prediction interval = [-1.222; 15.654], $I^2$ = 77.7\%).

Our last estimation analyses the relationship between $k$ and government
spending as a percentage of GDP. The coefficient is not statistically
significant, indicating a null effect (studies = `r mod[[7]]$k`, SMD = `r round(mod[[7]]$TE.random,3)`, SE = `r round(mod[[7]]$seTE.random,3)`, 95\% CI
= [`r round(mod[[7]]$lower.random,3)`; `r round(mod[[7]]$upper.random,3)`], $p$-value = `r round(mod[[7]]$pval.random,3)`, prediction interval = [-0.428; 0.423],
$I^2$ = 85.8\%). The result is very similar in the extended sample.

@@

In this article, we assess the empirical validity of the "law of $1/n$". Based
on a sample of `r length(unique(dat$id))` recent publications on the topic, our meta-analyses show that
there is little evidence that an increase in the number of legislators has a
significant effect on public expenditures. If such effect exists, it is likely
driven by an increase in $k$, the size of the upper legislature, as suggested
by several studies in the literature \citep[e.g.,][]{baqir2002districting,
bradbury2001legislative, bradbury2003local, chen_malhotra_2007,
gilligan2001fiscal, primo2006stop}. We find no robust evidence for the "reverse
law of $1/n$", which posits that larger legislatures lead to lower government
spending. One possible explanation is that the two logics are balancing each
other, thus leading us to a null impact of legislature size on public
expenditure.
